---
title: "Colon_Assignment1_Database"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Libraries 

```{r}

library(DBI)
library(tidyverse)
library(dplyr)
if(!require(readr)) install.packages("readr")
if(!require(skimr)) install.packages("skimr")
library(stringr)
library("RSQLite")


```

###1. Import and Tidy


```{r}

#create database
parcels_db <- dbConnect(RSQLite::SQLite(), dbname = "parcels.db")
parcels_db

#read in usecode file
UseCodes <- read_delim("UseCodes.csv", delim = "|", na = "NA",
                     col_types = cols_only(UseCode = "n", CodeDesc = "c"))

#read in parcel csv


parcels <-
  read_delim(
    "parcels.csv.zip",
    delim = "|",
    col_types = cols(.default = col_character())
  )




#load data into parcels db

dbWriteTable(parcels_db, "parcels", parcels, row.names = T, overwrite = T)
dbWriteTable(parcels_db, "UseCodes", UseCodes, row.names = T, overwrite = T)

#colnames(parcels)
#colnames(UseCodes)

```



###2. Analyze 

####What are the 10 most-frequently-occuring land uses (in descending order)?
*Hint: Load the UseCodes into a separate tibble, then join it to the parcels tibble.*

```{sql connection=parcels_db}

SELECT parcels.UseCode, UseCodes.CodeDesc, count(parcels.UseCode) AS "Count"
FROM parcels 
JOIN UseCodes ON parcels.UseCode = UseCodes.UseCode
Group By parcels.UseCode
ORDER BY Count DESC
Limit 10;

```


####How many acres are in agricultural preserves?
*Assume agricultural preserves are indicated by non-NA values for AgPres.*

```{sql connection=parcels_db}

Select sum(Acre) As "Total Acreage" 
From parcels
where Ag_Preserve_Num IS NOT NULL;


```



####What is the mean net assessed value per acre of the entire county?
*i.e., as if the entire county were a single parcel.*

```{sql connection=parcels_db}

Select (sum(parcels.Total_Value))/(sum(parcels.Acre)) AS "Mean Total Value"
FROM parcels;

```


####What is the total net assessed value of all non-taxable parcels?
*Assume non-taxable parcels are indicated by non-NA values for NonTaxCode.*

```{sql connection=parcels_db}

Select sum(parcels.Total_Value) As "Total Value for Non-taxable" 
From parcels
where parcels.Non_Tax_Property IS NOT NULL;

```


####What are the 10 largest property holders, by acreage?
*Use the complete mailing address for tax statements as a proxy for the parcel owner. Ignore parcels where either component of the address is NA.*

```{sql connection=parcels_db}

SELECT parcels.Address1, sum(parcels.Acre) AS "Total Property Acreage"
FROM parcels 
Where parcels.Address1 IS NOT NULL AND parcels.Acre IS NOT NULL
Group By parcels.Address1
ORDER BY "Total Property Acreage" DESC
Limit 10;

```


####What are the 10 largest property holders, by net assessed value?

```{sql connection=parcels_db}

SELECT parcels.Address1, sum(parcels.Total_Value) AS "Total Property Value"
FROM parcels 
Where parcels.Address1 IS NOT NULL AND parcels.Total_Value IS NOT NULL
Group By parcels.Address1
ORDER BY "Total Property Value" DESC
Limit 10;

```


