---
title: "ESM262_Assignment1_Data Wrangling"
author: "Mario Colon"
date: "4/30/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###1. Import and Tidy


```{r message=FALSE, warning = FALSE}

library(tidyverse)

#Read in raw data

parcels_raw <-
  read_delim(
    "Santa_Barbara_County_parcels_2011.csv.zip",
    delim = ",",
    col_types = cols(.default = col_character())
  )

#str(parcels_raw)

#APN (Chr), Situs1 (Chr), Situs2 (Chr), Acreage(Num), UseCode (Chr), NonTaxCode (Chr), AgPres (Chr), LandValue (Num), Net_Impr (Num), Net_AV (Num), M_Address1 (Chr), M_Address2 (Chr)

#Copy and convert 

parcels <- transmute(parcels_raw,
                    Parcel_Number = parse_character(APN, na = "Unknown"),
                    Short_Address = parse_character(Situs1, na = "Unknown"),
                    Long_Address = parse_character(Situs2, na = "Unknown"),
                    Acre = parse_number(Acreage, na = "0"),
                    UseCode = parse_character(UseCode, na = "Unknown"),
                    Non_Tax_Property = parse_character(NonTaxCode),
                    Ag_Preserve_Num = parse_character(AgPres), 
                    Land_Value = parse_number(LandValue, na = "0"), 
                    Improvement_Value = parse_number(Net_Impr, na = "0"), 
                    Total_Value = parse_number(Net_AV),
                    Address1 = parse_character(M_Address1, na = "Unknown"), 
                    Address2 = parse_character(M_Address2, na = "Unknown")
                    )

#str(parcels)

#Write csv file 

parcels <- write_delim(parcels,
  "parcels.csv",
  delim = "|",
  na = ""
)


```

###2. Analyze 

####What are the 10 most-frequently-occuring land uses (in descending order)?
*Hint: Load the UseCodes into a separate tibble, then join it to the parcels tibble.*


```{r message=FALSE, warning = FALSE}
#read in UseCodes file
use_code <- read_delim("UseCodes.csv", delim = "|")

#subset and merge with UseCode to determine 10 most-frequently-occuring land
land_use <- parcels %>% 
  left_join(use_code, by = "UseCode") %>% 
  filter(!is.na(UseCode)) %>% 
  group_by(UseCode, CodeDesc) %>% 
  select(UseCode, CodeDesc) %>% 
  count(UseCode) %>% 
  arrange(desc(n)) %>% 
  head(n = 10)

land_use  

```

####How many acres are in agricultural preserves?
*Assume agricultural preserves are indicated by non-NA values for AgPres.*

```{r message=FALSE, warning = FALSE}
#subset based on Ag_Preserve_Num, remove NA for that number to remove non-agricultural preserve parcels 
ag_preserve <- parcels %>% 
  left_join(use_code, by = "UseCode") %>%
  filter(Ag_Preserve_Num != "NA") %>%
  filter(Acre != "NA")
  
sum(ag_preserve$Acre) 

  
```


####What is the mean net assessed value per acre of the entire county?
*i.e., as if the entire county were a single parcel.*

```{r message=FALSE, warning = FALSE}

#subset Total_Value and Acre as two variables of interest
mean_net_value_county <- parcels %>% 
  left_join(use_code, by = "UseCode") %>%
  select("Total_Value", "Acre") %>%
  filter(Acre >=0)

#sum of average total value and acre
total_value_sum <- sum(mean_net_value_county$Total_Value)
acre_sum <- sum(mean_net_value_county$Acre)

#divide total value sum by sum of acres
mean_value_acre <- total_value_sum/acre_sum
mean_value_acre

```

####What is the total net assessed value of all non-taxable parcels?
*Assume non-taxable parcels are indicated by non-NA values for NonTaxCode.*

```{r message=FALSE, warning = FALSE}

#remove parcel that non-taxable
non_tax <-  parcels %>% 
  filter(!is.na(Non_Tax_Property)) 

#sum value of remaining parcels 
sum(non_tax$Total_Value)

```


####What are the 10 largest property holders, by acreage?
*Use the complete mailing address for tax statements as a proxy for the parcel owner. Ignore parcels where either component of the address is NA.*

```{r message=FALSE, warning = FALSE}

#Keep only parcels that have full 'M_ADDRESS1' and keep top 10 based on acre size
prop_owner_acre <- parcels %>% 
  select(Address1, Acre) %>% 
  filter(!is.na(Address1)) %>% 
  group_by(Address1) %>% 
  summarize(acre_size = sum(Acre)) %>% 
  arrange(desc(acre_size)) %>% 
  head(n = 10)

prop_owner_acre

```


####What are the 10 largest property holders, by net assessed value?

```{r message=FALSE, warning = FALSE}

#Keep only parcels that have full 'M_ADDRESS1' and keep top 10 based on property value
prop_owner_net_av <- parcels %>% 
  select(Address1, Total_Value) %>%  
  filter(!is.na(Address1)) %>% 
  group_by(Address1) %>% 
  summarize(total_value = sum(Total_Value)) %>% 
  arrange(desc(total_value)) %>% 
  head(n = 10)

prop_owner_net_av

```

