---
title: "Assignment3"
author: "Mario Colon"
date: "5/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

#Call in function 

source("autopower.R")


```


What if we design new highways that have lower rolling coefficients where we reduce the rolling coefficient by 50%

```{r}

#Adjust default value for crolling in the autopower() function from .015 *.5
crolling_50 = 0.0075 

#set nsamples
nsample = 1000

#Create dataframe for output of autopower function 
speeds = rnorm(mean=25, sd=4, n=nsample)
results_50 = data.frame(speed=speeds, power=NA)

##Generate 3 possible cars with respectvie area and mass
possible_cars_50 = data.frame(mass=c(31000,45000,38000), area = c(25,30,22))
possible_cars_50$row = seq(from=1, to=nrow(possible_cars_50))
possible_cars_50$prob = c(0.4, 0.4, 0.2)
##Created a large random sample of the three cars   
whichcar_50 = base::sample(possible_cars_50$row, size=nsample, prob=possible_cars_50$prob, replace=TRUE)

## Populate results_ms DF with the different masses and areas of the three cars.
results_50$mass = possible_cars_50$mass[whichcar_50]
results_50$area = possible_cars_50$area[whichcar_50]

#run function

pb_50_crolling <- autopower(V=results_50$speed, m = results_50$mass, A = results_50$area, crolling = crolling_50)


```



Alternatively, we implement a policy that reduce the mean speed from 100 km/h to 80 km/h (still with 10% standard deviation)

```{r}

#Reduce the mean speed from 100 km/h to 80 km/h (still with 10% standard deviation)

mean_speed_high = 100 * 0.277    #100km/hr to m/s
mean_speed_low = 80 * 0.277  #80km/hr to ms

set.seed(25)
speeds_1 = rnorm(mean = mean_speed_high, sd=0.1*mean_speed_high, n=nsample)
speeds_2 = rnorm(mean = mean_speed_low, sd=0.1*mean_speed_low, n=nsample)

#create database for reduced mean speed

#Create dataframe for output of autopower function 

results_2 = data.frame(speed=speeds_2, power=NA)

##Generate 3 possible cars with respectvie area and mass
possible_cars_2 = data.frame(mass=c(31000,4200,38000), area = c(25,30,22))
possible_cars_2$row = seq(from=1, to=nrow(possible_cars_2))
possible_cars_2$prob = c(0.4, 0.4, 0.2)
##Created a large random sample of the three cars   
whichcar_2 = base::sample(possible_cars_2$row, size=nsample, prob=possible_cars_2$prob, replace=TRUE)

## Populate results_ms DF with the different masses and areas of the three cars.
results_2$mass = possible_cars_2$mass[whichcar_2]
results_2$area = possible_cars_2$area[whichcar_2]

#run function

pb_2_speed <- autopower(V=results_2$speed, m = results_2$mass, A = results_2$area)


```


```{r}
#final comparison 

##mean power generation with lower crolling
mean(pb_50_crolling)

##mean power generation with lower speed 
mean(pb_2_speed)

```

Given the two potential changes in approaches (road design change vs reduced speed policy), it would seem that a policy of reduced speed to 80km/hr instead of increasing friction as result of road design change is better to reduce energy output by approximately 26,840.80 W.  

