---
title: "climate_analysis"
author: "Mario Colon"
date: "June 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load packages

```{r}

library(tidyverse)
library(dplyr)

```


#load in data

```{r}

climate <- read_delim("clim.txt", delim = " ")
climate
climate <- select(climate, month, day, year, tmax_c, tmin_c, precip)

climate

```
#Summarization Functions (two different functions that summarize climate data in some way (your choice - it could be finding a trend, or finding extremes; or even simply mean winter precipitation))

```{r}

#extreme temperatures

source("/extreme_temp.R")

extremes <- extreme_temp(climate$tmax_c, climate$tmin_c, climate$year)

extremes
  
```
```{r}

#mean winter precip 

mean_precip <- mean_precip(climate$precip, climate$year)
  
mean_precip

```

#Calculation Functions (two function that calculate some measure of impact (you can really make these up) -so for example you could create a function that computes a heat stress index based on air temperature with parameters that vary with for different species; or a function that estimates water supply from precipitation with parameters that vary to reflect water storage …you can make something up just as long as its somewhat sensible

```{r}

```


#Cost Function (one function that estimates the costs of impacts)
use multi-dimensional arrays in one of your functions

```{r}

```


